From Freebase to Wikidata: The Great Migration

### 2015-10-08 paper.tex
Reworked Section 3.3.

The quality of the data in Freebase was challenged by the Wikidata community.
For example, a~small city in France had an ISO country code%
\footnote{See \url{http://www.freebase.com/m/03cc0d0}}.
Another example, Boston (the city in Massachusetts) has the type \texttt{/people/person}.
A fully automatic upload of all the content from Freebase into Wikidata
did not seem advisable as the expectations of the community regarding
the quality of the content automatically added to Wikidata are high (and rightly so).
Also, whereas an upload of all the data might have lead to accelerated growth
in the short term, such a~fast upload might also endanger the sense of ownership
the Wikidata community has for its data.
Instead we decided to rely on human curation and created Primary Sources,
a simple tool which displays statements to the user that may be added to the currently shown item.
The user can, with only one click, reject the statement or approve it and save it to Wikidata.
It is currently available as a~gadget, an opt-in option in Wikidata.
This tool has no dependencies on Freebase and may be easily reused by other projects
which may want to import data into Wikidata.
For that we have implemented an upload API and a~feature that allows users
to display only statements from a~given dataset.
In fact, during the project one other researcher approached us in order to upload
a dataset extracted from natural language text.
A statistics dashboard has also been created in order to monitor the curation work
\url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}
and the source code of the tool is available on GitHub:
\url{https://github.com/google/primarysources}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/primary-sources.png}
\caption{Primary Sources user interface}
\end{figure}

Currently (August~18, 2015) the tool has been used by more than a~hundred users
who have performed more than 15,000 approve or reject actions.
More than 12M~statements are uploaded to the tool.

>>>>

The data quality of Freebase was examined thoroughly by the Wikidata community
and anecdotal evidence indeed confirmed that there were certain challenges.
For example, a~small city in France was assigned an ISO country code,%
\footnote{See \url{http://www.freebase.com/m/03cc0d0}}
then Boston (the city in Massachusetts) had the type \texttt{/people/person},
and several other occurrences of data quality issues were revealed.
In consequence, a~fully automatic upload of all the contents from Freebase into Wikidata
did not seem advisable as the expectations of the Wikidata community regarding the
quality of automatically uploaded data are high (and rightly so).
Despite the fact that an automated upload of all the data
would have led to accelerated growth in the short term,
such a~fast increase of available data would certainly have endangered
the sense of ownership of the Wikidata community for its data.
As an alternative approach, we decided to rely on crowdsourced human curation
and created the \emph{Primary Sources Tool},
a~software solution that displays Freebase statements
for verification to the user that can be added to the currently shown Wikidata item.
The tool will be described in more detail in \autoref{sec:primary-sources-tool},
we provide a~short summary of its functionality here.
\autoref{fig:primary-sources-tool} shows a~screenshot of the user interface
of the tool that integrates nicely into the Wikidata user interface.
The user can, with just one click, reject or approve a~statement,
and, in case of approval, therewith add the statement to Wikidata.
The Primary Sources Tool is currently available as a~so-called gadget,
\emph{i.e.}, as an opt-in feature in Wikidata.
This tool has no dependencies on Freebase and can be easily reused for other projects
that want to import datasets into Wikidata.
Therefore, we have implemented an upload API in the back-end
and a~filter in the front-end that allows users to display only statements from a~given dataset.
This flexibility is much-wanted, and during the development period of the project
another researcher approached us in order to upload a~dataset extracted from natural language text.
To visualize the migration progress, we have created a~realtime dashboard
available at \url{https://tools.wmflabs.org/wikidata-primary-sources/status.html}.
The code of the Primary Sources Tool is openly available
(\url{https://github.com/google/primarysources}) under the terms of the Apache~2.0 license.
At the time of writing (October, 2015), the tool has been used by more than hundred users
who performed about 30,000 approval or rejection actions.
More than 12M~statements have been uploaded in total.

\begin{figure}[!htbp]
\centering
\includegraphics[width=8.45 cm]{img/primary-sources.png}
\caption{Primary Sources Tool user interface}
\label{fig:primary-sources-tool}
\end{figure}
